What\,Agreement\,Hath\,Mars\,With\,Earth? \\ An Exploration of Urbit's IPC Interfaces

### 2024-04-10 mss.tex
Proof

%\newcommand{\authorname}{}
\newcommand{\authorpatp}{\patp{mopfel-winrux}}
\newcommand{\affiliation}{Native Planet}

%  Make first page footer:
\fancypagestyle{firststyle}{%
\fancyhf{}% Clear header/footer
\fancyhead{}
\fancyfoot[L]{{\footnotesize
              %% We toggle between these:
              % Manuscript submitted for review.\\
              {\it Urbit Systems Technical Journal} I:1 (2024):  1–3. \\
              ~ \\
              Address author correspondence to \authorpatp.
              }}
}
%  Arrange subsequent pages:
\fancyhf{}
\fancyhead[LE]{{\urbitfont Urbit Systems Technical Journal}}
\fancyhead[RO]{Urbit's {\sc ipc} Interfaces}
\fancyfoot[LE,RO]{\thepage}

@@

% We will adjust page numbering in final editing.
\pagenumbering{arabic}
\setcounter{page}{1}

@@

Arvo's unique architecture avoids what is referred to as ``event spaghetti'' by maintaining a clear causal chain for every computation \citep{Whitepaper}.  Each chain begins with a Unix I/O event and progresses through a series of steps until the computation either completes successfully or fails.  This deterministic nature allows for a high level of predictability and control, a stark contrast to most Earth-based operating systems.

@@

Communication between the \texttt{king} and \texttt{serf} is achieved via an inter-process communication ({\sc ipc}) port.  This {\sc ipc} mechanism facilitates a continuous dialogue between the two components, ensuring that the \texttt{king} can effectively manage external interactions while the \texttt{serf} maintains the integrity of the Arvo VM.

@@

The \texttt{\%khan} vane functions as a high-level thread interface for managing both internal and external communications within the Urbit ecosystem.  \texttt{\%khan} is designed to efficiently handle complex I/O operations, employing an I/O monad coupled with an exception monad.  This approach allows \texttt{\%khan} to adeptly manage complex I/O tasks while effectively handling potential failures.

@@

\sloppy
The \texttt{\%khan} vane was conceived as a way to control Urbit ships from the exterior using threads.\footnote{\texttt{\%khan} was originally called the ``control plane'' after the network routing distinction between the control plane and the data plane.}  \texttt{\%khan}'s conception evolved a fair bit from proposal to implementation.  In practice, \texttt{\%khan} is essentially an interface wrapper for \texttt{\%spider}-based threads,\footnote{\texttt{\%spider} is an agent for managing transient thread-level operations, i.e. in \texttt{\%gall}.} which produces the topsy-turvy situation in which a vane relies on a piece of userspace infrastructure to function correctly.  \texttt{\%khan} allows for pre-written threads that allow for easy hosting and maintenance to be bundled and distributed, helping both hosting companies and self-hosters to manage Urbit instances efficiently.

@@

The \texttt{conn.c} driver in Vere\footnote{\href{https://github.com/urbit/vere/blob/develop/pkg/vere/io/conn.c}{\texttt{vere/io/conn.c}}} forms part of the \texttt{king} process.  It exposes a \href{https://en.wikipedia.org/wiki/Unix_domain_socket}{Unix domain socket} at \texttt{/path/to/pier/.urb/conn.sock} for sending and receiving data from external processes.

>>>>

%\newcommand{\authorname}{}
\newcommand{\authorpatp}{\patp{mopfel-winrux}}
\newcommand{\affiliation}{NativePlanet}

%  Make first page footer:
\fancypagestyle{firststyle}{%
\fancyhf{}% Clear header/footer
\fancyhead{}
\fancyfoot[L]{{\footnotesize
              %% We toggle between these:
              % Manuscript submitted for review.\\
              {\it Urbit Systems Technical Journal} I:1 (2024):  83–92. \\
              ~ \\
              Address author correspondence to \authorpatp.
              }}
}
%  Arrange subsequent pages:
\fancyhf{}
\fancyhead[LE]{{\urbitfont Urbit Systems Technical Journal}}
\fancyhead[RO]{Urbit's {\sc ipc} Interfaces}
\fancyfoot[LE,RO]{\thepage}

@@

% We will adjust page numbering in final editing.
\pagenumbering{arabic}
\setcounter{page}{83}

@@

Arvo's unique architecture avoids what is referred to as ``event spaghetti'' by maintaining a clear causal chain for every computation \citep{Whitepaper}.  Each chain begins with a Unix I/O event and progresses through a series of steps until the computation completes successfully or fails.  This deterministic nature allows for a high degree of predictability, a stark contrast to most Earth-based operating systems.

@@

Communication between the \texttt{king} and \texttt{serf} is achieved via an inter-process communication ({\sc ipc}) port.  This {\sc ipc} mechanism facilitates a dialogue between the two components, ensuring that the \texttt{king} can effectively manage external interactions while the \texttt{serf} maintains the integrity of the Arvo VM.

@@

The \texttt{\%khan} vane functions as a high-level thread interface for managing both internal and external communications within the Urbit ecosystem.  \texttt{\%khan} is designed to efficiently handle complex I/O operations, employing an I/O monad coupled with an exception monad.  This approach allows \texttt{\%khan} to manage complex I/O tasks while robustly handling potential failures.

@@

\sloppy
The \texttt{\%khan} vane was conceived as a way to control Urbit ships from the exterior using threads.\footnote{\texttt{\%khan} was originally called the ``control plane'' after the network routing distinction between the control plane and the data plane.}  \texttt{\%khan}'s conception evolved a fair bit from proposal to implementation.  In practice, \texttt{\%khan} is essentially an interface wrapper for \texttt{\%spider}-based threads,\footnote{\texttt{\%spider} is a \texttt{\%gall} agent to manage transient thread-level operations.} which produces the topsy-turvy situation in which a vane relies on a piece of userspace infrastructure to function correctly.  \texttt{\%khan} allows for pre-written threads that allow for easy hosting and maintenance to be bundled and distributed, helping both hosting companies and self-hosters to manage Urbit instances efficiently.

@@

The \texttt{conn.c} driver in Vere\footnote{\href{https://github.com/urbit/vere/blob/develop/pkg/vere/io/conn.c}{\texttt{vere/io/conn.c}}} forms part of the \texttt{king} process.  It exposes a \href{https://en.wikipedia.org/wiki/Unix_domain_socket}{Unix domain socket} at \texttt{/path/to/pier} \texttt{/.urb/conn.sock} for sending and receiving data from external processes.
